{\bf Demystifying Debugging}

### 2012-11-07 gpi-debug/debugging.tex
code-complete lecture slides

\begin{frame}{``The Universal Backup Plan''}
	\textbf{15-112, 15-122, 15-150, ...}
	\begin{itemize}
		\item What do these classes all have in common?
		\pause
		\item How to write code that {\em works}
	\end{itemize}
	\pause
	{\bf 98-172}
	\begin{itemize}
		\item What to do with code that {\em doesn't}
	\end{itemize}
	% TODO: a picture goes here -- blank slate vs equipped/prepared technician
\end{frame}

\begin{frame}{Take-away}
	\textbf{Debugging toolbox}
	\begin{itemize}
		\item Two major {\em strategies} for approaching bugs
		\item Assorted {\em tactics} for applying to broken programs
	\end{itemize}
	% Say: "It's not, e.g., a checklist of bugs that you can just run down and test one-by-one; it's more abstract than that. You have to use your brain to figure out how to apply each tool. But the hope is that you won't feel lost staring at a blank slate of possible approaches.
\end{frame}

@@

\begin{frame}{Aside: Asking for Help}
	This DOESN'T mean...
	\begin{itemize}
		\item That TAs can't help you without a detailed story
		\item (That you shouldn't seek help until you've already found your bug?!)
	\end{itemize}
	\pause
	This does mean...
	\begin{itemize}
		\item You need a detailed story to understand your bug.
		\item If the TA doesn't know the story, how can they help?
		\pause
		\begin{itemize}
			\item They know what you should do to make your story better.
		\end{itemize}
	\end{itemize}
\end{frame}

@@

\begin{frame}{Print Debugging}
	New story:

@@

\begin{frame}{Asserts: Stealthier, more concealable prints}
	Adding print statements everywhere is very verbose
	\begin{itemize}
		\item Pro: Get to see everything
		\item Con: Have to look at everything
	\end{itemize}
	\pause
	\linegap

	% TODO
\end{frame}

@@

\begin{frame}{Minimizing the Test Case}
	New story:

@@

% TODO
%\begin{frame}

>>>>

\begin{frame}{``The Universal Backup Plan''}
	\textbf{15-112, 15-122, 15-150, ...}
	\begin{itemize}
		\item What do these classes all have in common?
		\pause
		\item How to write code that {\em works}
	\end{itemize}
	\pause
	\linegap

	{\bf 98-172}
	\begin{itemize}
		\item What to do with code that {\em doesn't}
	\end{itemize}
	% TODO: a picture goes here -- blank slate vs equipped/prepared technician
\end{frame}

\begin{frame}{Take-away}
	This lecture's goal: To give you a {\em framework} for thinking when faced with buggy code.
	\linegap

	\textbf{Debugging toolbox}
	\begin{itemize}
		\item Two major {\em strategies} for approaching bugs
		\item Assorted {\em tactics} for applying to broken programs
	\end{itemize}
	% Say: "It's not, e.g., a checklist of bugs that you can just run down and test one-by-one; it's more abstract than that. You have to use your brain to figure out how to apply each tool. But the hope is that you won't feel lost staring at a blank slate of possible approaches.
\end{frame}

@@

\begin{frame}{Aside: Asking for Help}
	This DOESN'T mean...
	\begin{itemize}
		\item That TAs can't help you without a detailed story
		\item (That you shouldn't seek help until you've already found your bug?!)
	\end{itemize}
	\pause
	\linegap

	This does mean...
	\begin{itemize}
		\item You need a detailed story to understand your bug.
		\item If the TA doesn't know the story, how can they help?
		\pause
		\begin{itemize}
			\item They know what you should do to make your story better.
		\end{itemize}
	\end{itemize}
\end{frame}

@@

\begin{frame}{Print Debugging}
	Old story:

	\includegraphics[width=\textwidth]{time0.png}
\end{frame}
\begin{frame}{Print Debugging}
	New story:

@@

\begin{frame}{Asserts: Stealthier, more concealable prints}
	Adding print statements everywhere is very verbose
	\begin{itemize}
		\item Pro: Get to see everything
		\item Con: Have to look at everything
	\end{itemize}
	\pause
	15-122's C0 ``contracts'' are assertions
	\begin{itemize}
		\item You don't need to use C0 to write assert-ful code
		\item Asserts can go anywhere, not just around functions/loops
	\end{itemize}
	\pause
	\linegap

	\texttt{~~~~\hilight{brown}{if}~(something\_that\_should\_never\_happen)~\{} \\
	\texttt{~~~~~~~~print(\hilight{brickred}{"it~happened!"});} \\
	\texttt{~~~~~~~~crash();} \\
	\texttt{~~~~\}} \\
\end{frame}
\begin{frame}{Asserts: Stealthier, more concealable prints}
	Adding print statements everywhere is very verbose
	\begin{itemize}
		\item Pro: Get to see everything
		\item Con: Have to look at everything
	\end{itemize}
	15-122's C0 ``contracts'' are assertions
	\begin{itemize}
		\item You don't need to use C0 to write assert-ful code
		\item Asserts can go anywhere, not just around functions/loops
	\end{itemize}
	\linegap

		\texttt{~~~~\hilight{brown}{foreach}~element~\hilight{brown}{in}~input~\{} \\
		\texttt{~~~~~~~~insert\_sorted(output,~element);} \\
		\texttt{~~~~~~~~\hilight{olivegreen}{assert(is\_sorted(output));}} \\
		\texttt{~~~~\}} \\
\end{frame}

@@

\begin{frame}{Minimizing the Test Case}
	Old story:

	\begin{center}
	\includegraphics[width=0.65\textwidth]{space0.png}
	\end{center}
\end{frame}
\begin{frame}{Minimizing the Test Case}
	New story:

@@

% TODO
\begin{frame}{Rubber Duck / Cardboard Dog Debugging}
	\begin{columns}[l]
	\column{0.75\textwidth}
	Hacker folklore says: describe your problem to a duck!
	\begin{itemize}
		\item ``What could I possibly get out of talking to an inanimate object?''
	\end{itemize}
	\linegap

	\onslide<2->{
	Hearing yourself speak makes you {\bf question your assumptions}.
	\begin{itemize}
		\item ``The bug can't be in X.''
		\item ``Why not? Are you sure X works?''
	\end{itemize}
	\linegap

	Helps figure out what part of the story to start searching at
	}
	\column{0.25\textwidth}
	\includegraphics[width=\textwidth]{rubberduck.jpg}
	\end{columns}
\end{frame}

\begin{frame}{Draw Pictures}
	Draw what you think should happen

	Label relevant variables and their values
	\linegap

	\begin{center}
		\includegraphics[width=0.8\textwidth]{draw-a-picture.png}
	\end{center}
\end{frame}
